---
title: "Homework 3: More Data Viz"
execute: 
  warning: true
fig-height: 3
fig-width: 5
fig-align: center
code-fold: false
---


We will be using the following packages.  If any is not installed on your machine, install it by either 

a. running the command `install.packages(package name)` in the Console or 
b. via the Packages panel --> Install --> type `package name` --> click Install.

```{r}
# Load required libraries
library(tidyverse)
library(openintro)
library(leaflet)
```

## Spatial Viz

### Exercise 1

In this exercise, we will use the same Starbucks location data we worked with in class but on the US state-level only while taking into account the population of each state.



```{r}
# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")
```

The code below creates the dataset `starbucks_us_by_state` that gives the number of Starbucks in each state. 

```{r}
starbucks_us_by_state <- starbucks |>
  filter(Country == "US") |>
  count(State.Province) |>
  mutate(state_name = str_to_lower(abbr2state(State.Province)))
```

The code below adds the variable `starbucks_per_10000` that gives the number of Starbucks per 10,000 people to the dataest `starbucks_with_2018_pop_est` which will be used for the spatial visualization.

```{r}
census_pop_est_2018 <- read_csv("https://mac-stat.github.io/data/us_census_2018_state_pop_est.csv") |>
  separate(state, into = c("dot", "state"), extra = "merge") |>
  select(-dot) |>
  mutate(state = str_to_lower(state))

starbucks_with_2018_pop_est <-
  starbucks_us_by_state |>
  left_join(census_pop_est_2018,
    by = c("state_name" = "state")
  ) |>
  mutate(starbucks_per_10000 = (n / est_pop_2018) * 10000)
```

#### Part a

Create a choropleth state map that shows the number of Starbucks per 10,000 people on a map of the US while taking into consideration the following instructions:

- Use a new fill color palette for the states,
- Add points for all Starbucks in the contiguous US,
- Add an informative title for the plot, and
- Include a caption that says who created the plot (you!).

```{r}
library(ggthemes)

states_map <- map_data("state")
```


```{r}
# 
ggplot(starbucks_with_2018_pop_est) + 
  geom_map(map= states_map, aes(map_id = state_name, fill= starbucks_per_10000)) +
  geom_point(data = starbucks %>% filter(Country=="US"),
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE, color="orange") + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()
```

#### Part b

Make a conclusion about what you observe from that spatial visual.

::: {.callout-note title="Answer"}

MN Map:

:::


### Exercise 2

In this exercise, you are going to create a single Leaflet map of some of your favorite places! The end result will be one map.

#### Part a

Create a data set using the tribble() function that has 10-15 rows of your favorite places. The columns will be the name of the location, the latitude, the longitude, and a column that indicates if it is in your top 3 favorite locations or not. For an example of how to use tribble(), look at the `favorite_stp` that is created manually below.

```{r eval=FALSE}
# Brianna's favorite St. Paul places - Used Google Maps to get coordinates
# https://support.google.com/maps/answer/18539?hl=en&co=GENIE.Platform%3DDesktop
favorite_stp <- tibble(
  ~place, ~long, ~lat, ~favorite,
  "Macalester College", -93.1712321, 44.9378965, "yes", 
  "Groveland Recreation Center", -93.1851310, 44.9351034, "yes",
  "Due Focacceria", -93.1775469,  44.9274973, "yes",
  "Shadow Falls Park", -93.1944518, 44.9433359, "no",
  "Mattocks Park", -93.171057, 44.9284142, "no",
  "Carondelet Fields", -93.1582673, 44.9251236, "no",
  "Pizza Luce", -93.1524256, 44.9468848, "no",
  "Cold Front Ice Cream", -93.156652, 44.9266768, "no"
)
```

```{r}
#suweda's favorite places

suweda_places <- tibble(
  place = c("Winchell Trail", "BuckHill Tubing",  "The Bean", "Cafe Astoria", "Como Zoo","Macalester College", "Guthrie Theater", "Central Park Zoo", "Botanical Garden","Niagra Falls"),
  lat= c(44.924012383445884,
          44.724935444213024,
          41.8827,
          44.94150072716086,
          44.981819836566636, 
          44.93813570998666,
          44.97844328590848, 
          40.768597162622356,
          33.790135103799855,
          43.09404510718638 ),
  long = c(-93.20430035960858,
          -93.28322299268974,
          -87.6233,-93.10657028645764,
          -93.15059996664003,-93.16672, -93.25391, -73.96228,
          -84.37262,
          -79.02666),
  favorite = c("no","yes","no","no","no","no","no","no","yes","yes")
    )
  
```

#### Part b

Create a map that uses circles to indicate your favorite places while taking into consideration the following instructions:

- Label the circles with the name of the place. 
- Choose the base map you like best. 
- Color your 3 favorite places differently than the ones that are not in your top 3. 
- Add a legend that explains what the colors mean

```{r}
library(gplots)
library(RColorBrewer)

pal <-  colorFactor(palette = c("blue", "green"), domain = suweda_places$favorite) 

leaflet(data=suweda_places)|> 
  addTiles() |>
  addCircles(color = ~pal(favorite), label = ~place) |> addLegend(pal = pal, values = ~favorite, group = "circles", position = "bottomleft")


```




## TidyTuesday

Tidy Tuesday is a weekly data project put on by some folks from the R Data Science community. Each week, a different data set is posted and people around the world wrangle and visualize that data. According to the organizers, "The intent of Tidy Tuesday is to provide a safe and supportive forum for individuals to practice their wrangling and data visualization skills independent of drawing conclusions."


The goals of this TidyTuesday are:

1.  Practice generating questions. You have to decide what to ask and how to answer it with a graphic.
2.  Practice identifying what viz and (eventually) wrangling tools are useful for addressing your questions.
4.  Get a sense of the broader data science community. Check out what people share out on [X/Twitter #TidyTuesday](https://x.com/search?q=%23TidyTuesday). Maybe even share your own #TidyTuesday work on social media. Recent Mac alum [Erin Franke (X/Twitter)](https://x.com/efranke7282) has an inspiring account! Scrolling through, you'll notice the trajectory of her work, starting from COMP/STAT 112 to today. Very cool.


### Exercise 3

Go to [TidyTuesday](https://github.com/rfordatascience/tidytuesday). Pick a dataset that was posted in **July, August, or September 2024**. Here, include:

- A short (\~2 sentence) written description of your data. This should include: 
  - the original data source, ie, where did TidyTuesday get the data from?
  - units of observation, ie, what are you analyzing?, and 
  - the data size, ie, how many data points do you have? how many variables are measured on each data point?.

- Code to import and examine the basic properties of your chosen dataset. This code must support the facts you cited in your short written description.



### Exercise 4


In the 3 sections below (Viz 1, 2, 3), construct 3 separate graphs that tell a *connected* story about the data you chose.  

Before each viz:

-   write a simple but specific research question you're trying to address with the viz.
-   write a 2-4 sentence summary of what you learn from the viz. This should connect back to your research question!

After each viz:

-   Comment on at least 2 effective aspects of the viz (consider the effective visualization principles).
-   Comment on at least 2 aspects of the visualization that could be improved. Perhaps these are aspects that you don't know how to implement yet but wish you could update it.

Make sure each viz:

-   has meaningful axis labels and legend titles
-   has a figure caption (fig.cap)
-   uses alt text (fig.alt)
-   uses a more color-blind friendly color palette


Tips:

-   Start with some questions in mind of what you want to learn.
-   Start with a simple viz (viz 1), and build this up into something multivariate (viz 3).
-   Reflect on each viz -- what new questions do you have after checking out the viz? Let these questions guide your next viz. (eg: recall how we worked through the `MacNaturalGas` data at the start of the Spatial Viz activity).


#### Viz 1

```{r eval=FALSE}
library(readr)
library(tidyverse)
plants <- ("/Users/suwedasaid/Documents/GitHub/mac-112-portfolio/homeworks/planting_2021.csv")

ggplot(plants, aes(x=number_seeds_planted)) + geom_histogram()

```

#### Viz 2

```{r eval=FALSE}
ggplot(plants, aes(x=vegetable, fill=variety)) + geom_bar()
```



#### Viz 3

```{r eval=FALSE}
ggplot(data=plants, aes(x=vegetable, y=number_seeds_planted)) + geom_boxplot()
```

